A Unified Treatment of Tractability for \\ Approximation Problems Defined on Hilbert Spaces

### 2024-03-12 Older_versions/GeneralizedTractability2023Sept21.tex
Updates from Overleaf

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)\footnote{We note that the radius of the $\cb_d$ could be taken to be an arbitrary positive value, $R$, in which case the problem becomes equivalent to the original one with the tolerance $\varepsilon$ replaced by $\varepsilon/R$ since the problem and the optimal approximate soluion are both linear (see below).}.
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

@@

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$, and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there focus mostly on relating the properties of the function $T$ to the different varieties of tractability, and not in terms of sums of (functions of) the singular values of a problem. Hence, the present paper adds to what is presently known on this subject. 
In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

@@

\begin{proof}
    \textbf{Sufficient condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge \boldsymbol{0}$.
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $\thed=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  +
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ]
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp)
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}} .
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

>>>>

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$, defined on the unit ball of functions in $\cf_d$, denoted $\cb_d$, which  satisfy an absolute error criterion:
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)\footnote{We note that the radius of the $\cb_d$ could be taken to be an arbitrary positive value, $R$, in which case the problem becomes equivalent to the original one with the tolerance $\varepsilon$ replaced by $\varepsilon/R$ since the problem and the optimal approximate solution are both linear (see below).}.
\end{equation}
Here, $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

@@

We want to bound the information complexity in terms of a simple function, $T$, \kachi{which depends of three variables} $\varepsilon^{-1}$, \kachi{ $d$ and some parameter $\vp$},  and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there focus mostly on relating the properties of the function $T$ to the different varieties of tractability, and not in terms of sums of (functions of) the singular values of a problem. Hence, the present paper adds to what is presently known on this subject. 
In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

@@

\begin{proof}
    \textbf{Sufficient condition:}\\
Suppose that \eqref{eq:tractiff4} holds for some $\vp \ge \boldsymbol{0}$.
By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $\thed=d$ and $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that
\begin{align*}
       \comp(\varepsilon,d)
       & \le  \kachi{T(\varepsilon^{-1},\thed,\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},\thed,\vp)}  +  \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \Biggr] \quad \forall  \thed, N \in \naturals}
       \\
      & \le T(\varepsilon^{-1},d,\vp) \Biggl [
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  +
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       \Biggr ]
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp)
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}} .
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.
